---
title: "Suicide Rates Final Project"
author: "Sydney Cindrich and Mancy Tomar"
date: "10/22/2020"
output:
  word_document: default
  html_document: default
---
```{r}
setwd("C:/Users/cindrich/Documents/Fall 2020/Classes/STAT 579")
suicide<-read.csv("master.csv")
```

##Getting familiar with the data
```{r}
str(suicide) #the data is either an integer or character
dim(suicide) #There are 43776 observation and 6 variables
sum(is.na(suicide)) #There are 7716 missing values in the data set
summary(suicide) #The missing values are from suicide_no (2256) and population (5460)
```
##Discover what country has the highest suicide rate
```{r}
library(dplyr)
suicide%>%
  mutate( 
   Rank= rank(-suicides_no)
  )%>%
  arrange(Rank)
```
##Russian Federation has the highest suicide rate

##Discover what country has the lowest suicide rate
```{r}
suicide%>%
  mutate( 
    Rank= rank(suicides_no)
  )%>%
  filter(!is.na(Rank))%>%
  arrange(Rank)
```
##Discover what year has the highest suicide rate? What year has the lowest?
##Is suicide increasing over the years? Decreasing? In what countries?
##What age is more likely to committ suicide? Is this constant throughout the years?
##What sex is most likely to committ suicide? Is this constant throughout the years
##How are country and age associated with suicide rates?
#How are country and sex associated with suicide rates
##What generation has the highest suicide rate?

#Introducing a new data set to see what state in the U.S. has the highest suicide rate
```{r}
setwd("C:/Users/cindrich/Documents/Fall 2020/Classes/STAT 579")
suicide.2<-read.csv("Suicide Mortality by State (1).csv")
summary(suicide.2)
head(suicide.2)
unique(suicide.2$YEAR)
```
##What state has the highest suicide rate in the U.S.? The lowest?
##What year has the highest suicide rate in the U.S.? The lowest?

